A general Nullstellensatz for generalized spaces

### 2019-03-17 paper-qcoh.tex
Give details how definability entails the Nullstellensatz

% An immediate corollary of Theorem~\ref{thm:definability} is the
% Nullstellensatz of Theorem~\ref{thm:nullstellensatz}. Indeed, arguing
% internally to~$\Set[\TT]$, let~$\sigma = (\varphi \seq{\vec x} \psi)$ be a
% geometric$^\star$ sequent over the signature of~$\ul{\TT}/U_\TT$. Assume
% that~$\sigma$ holds for~$U_\TT$. Then the subsets~$\{ (\vec x) \,|\, \varphi
% \}$ and~$\{ (\vec x) \,|\, \varphi \wedge \psi \}$ are equal. Hence, by
% Theorem~\ref{thm:definability}, the formulas~$\varphi$
% and~$\varphi \wedge \psi$ are provably$^\star$ equivalent.
% Thus~$\ul{\TT}/U_\TT$ proves$^\star$~$(\varphi \seq{\vec x} \psi)$.

@@

\begin{proof}Theorem~\ref{thm:definability} on the semantic side and the axiom
scheme~\eqref{eq:higher-order-nullstellensatz} on the syntactic side allow us to replace any mention of a
powersort~$P(X)$ in~$\alpha$
by~$\mathrm{Form}^\star_{x\!\?\!X}(\ul{\TT}/U_\TT)/({\dashv\vdash_x})$,
similarly to how the proof of Lemma~\ref{lemma:extended-conservative} compiles
extended geometric logic to ordinary geometric. Then we
can argue as in the proof of Theorem~\ref{thm:characterization}, noting that
the axiom scheme indeed entails the axiom scheme~\eqref{eq:nullstellensatz} posited by
Theorem~\ref{thm:characterization}.\end{proof}

>>>>

\begin{rem}\label{rem:definability-entails-nullstellensatz}
The proof of Theorem~\ref{thm:definability} used the Nullstellensatz
of Theorem~\ref{thm:nullstellensatz}. Conversely, the internal statement of
Theorem~\ref{thm:definability} entails the internal
statement~\eqref{eq:nullstellensatz} of Theorem~\ref{thm:nullstellensatz} in
extended geometric logic. Indeed, leaving the translation to formal logic to
the reader, let~$\sigma = (\varphi \seq{\vec x} \psi)$ be a
geometric$^\star$ sequent over the signature of~$\ul{\TT}/U_\TT$ such
that~$\sigma$ holds for~$U_\TT$. Then the subsets~$\{ (\vec x) \,|\, \varphi
\}$ and~$\{ (\vec x) \,|\, \varphi \wedge \psi \}$ are equal. Hence the
assumption implies that the formulas~$\varphi$
and~$\varphi \wedge \psi$ are provably$^\star$ equivalent.
Thus~$\ul{\TT}/U_\TT$ proves$^\star$~$(\varphi \seq{\vec x} \psi)$.
\end{rem}

@@

\begin{proof}Theorem~\ref{thm:definability} on the semantic side and the axiom
scheme~\eqref{eq:higher-order-nullstellensatz} on the syntactic side allow us to replace any mention of a
powersort~$P(X)$ in~$\alpha$
by~$\mathrm{Form}^\star_{x\!\?\!X}(\ul{\TT}/U_\TT)/({\dashv\vdash_x})$,
similarly to how the proof of Lemma~\ref{lemma:extended-conservative} compiles
extended geometric logic to ordinary geometric. Then we
can argue as in the proof of Theorem~\ref{thm:characterization}, noting that
the axiom scheme indeed entails the axiom scheme~\eqref{eq:nullstellensatz} by
Remark~\ref{rem:definability-entails-nullstellensatz}.\end{proof}
